---
title: "Whisky Data"
subtitle: "Information about 86 Scottish Single Malt Whiskies and Distilleries"
output:
  html_document:
    theme: cosmo
---

<style type="text/css">
body, td {
   font-family: EB Garamond;
   font-size: 18px
}
h1.title {
   font-family: EB Garamond Medium;
   font-size: 38px
}
h1 { /* Header 1 */
   font-family: EB Garamond Medium;
   font-size: 28px
}
h2 { /* Header 2 */
   font-family: EB Garamond Medium;
   font-size: 22px
}
h3 { /* Header 3 */
   font-family: EB Garamond Medium;
   font-size: 18px
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
```

# R packages and setup

```{r packages}
library(tidyverse)
library(stringr)
library(skimr)
library(ggrepel)
library(broom)
library(MASS)
library(corrr)
library(ggdendro)
```

# Obtain data

Just download the file *whiskies.txt* `if` it is non-existing in the local path.

```{r}
whisky_data <- "whiskies.txt"

if (! whisky_data %in% list.files()) {
   
   whisky_url <- "https://outreach.mathstat.strath.ac.uk/outreach/nessie/datasets/whiskies.txt"
   
   download.file(whisky_url, whisky_data)
   
}

whisky_data_raw <- read.delim("whiskies.txt", sep = ",")
```

# Data management and wrangling

Using different functions to get basic insights of the used whisky data.

```{r}
whisky_df <- whisky_data_raw %>%
   dplyr::select(! RowID) %>% 
   mutate(Postcode = str_trim(Postcode, side = "left")) 

whisky_df %>% 
  glimpse()

whisky_df %>% 
   head()

whisky_df %>% 
  skim()

# data in long format
whisky_df_long <- whisky_df %>% 
   pivot_longer(cols = Body:Floral, names_to = "taste_category", values_to = "value")

# summarising taste data
whisky_df_long %>% 
   group_by(taste_category) %>% 
   summarise(mean = mean(value),
             median = median(value))
```

# Statistical analysis

## Correlations

In order to analyze the relationship between the different taste categories the individual correlations (pearson $\rho$) are computed and displayed.

$$ \Large \rho = \frac{cov(X,Y)}{\sigma_X \sigma_Y}$$
```{r}
# Correlation matrix
whisky_df %>% 
   dplyr::select(Body:Floral) %>% 
   correlate() %>% 
   rearrange() %>%
   shave() %>% 
   fashion() 
   
# Correlation plot
corr_plot <- whisky_df %>% 
   dplyr::select(Body:Floral) %>% 
   correlate() %>% 
   rearrange() %>%
   shave() %>% 
   rplot() 

corr_plot

ggsave("corr_plot.png", 
       corr_plot, 
       width = 8, 
       height = 4)

# Correlation network
network_plot <- whisky_df %>% 
   dplyr::select(Body:Floral) %>% 
   correlate() %>% 
   network_plot()

network_plot

ggsave("network_plot.png", 
       network_plot, 
       width = 8, 
       height = 4)

```

## Dendrogram

### Dendogram for distilleries

```{r, fig.height = 10, fig.width=6}
# Dendrogram distilleries
whisky_df %>% 
   column_to_rownames("Distillery") %>% 
   dplyr::select(Body:Floral) %>% 
   dist() %>% 
   hclust() %>% 
   ggdendrogram(rotate = TRUE) + 
   labs(title = "Dendrogam | Distilleries") + 
   theme(text = element_text(family = "EB Garamond"),
         plot.title = element_text(size = 20)) +
   ggsave("dendrogram_dist.png", 
          height = 10,
          width = 6)
```

### Dendrogram for taste categories

```{r, fig.height = 10, fig.width=6}
# Dendrogram taste categories
whisky_df %>% 
   dplyr::select(Distillery, Body:Floral) %>% 
   pivot_longer(cols = Body:Floral, names_to = "taste_category", values_to = "taste_value") %>% 
   pivot_wider(names_from = Distillery, values_from = taste_value) %>% 
   column_to_rownames("taste_category") %>% 
   dist() %>% 
   hclust() %>% 
   ggdendrogram(rotate = TRUE) +
   labs(title = "Dendrogam | Taste Categories") + 
   theme(text = element_text(family = "EB Garamond"),
         plot.title = element_text(size = 20)) +
   ggsave("dendrogram_taste.png", 
          height = 7.5,
          width = 4.5)
```

## *k*-means clustering

The *k*-means algorithm is used to create distinct clusters among the distilleries based on the values of the tasting categories.


## Linear regression / linear model

```{r, include = FALSE}
# model_data <- whisky_df %>% 
#    transmute(across(c(Body:Floral), as.factor))






whisky_df %>% 
   dplyr::select(Body:Floral)%>% 
   map(~ lm(Body ~ .x, data = whisky_df)) %>% 
   map(summary) %>% 
   map_dbl("r.squared") %>% 
   tidy %>% 
   arrange(desc(x)) %>% 
   rename(r.squared = x) %>% 
   knitr::kable()


# model_data <- whisky_df %>% 
#    mutate(across(c(Body:Floral), as.factor)) %>% 
#    mutate(Longitude = as.numeric(str_trim(as.character(Longitude)),
#           Latitude = as.numeric(str_trim(as.character(Latitude))))) %>% 
#    filter(is.infinite(Latitude))
   # mutate(Body = factor(Body, levels = 0:4, ordered = TRUE))
          # Body_chr = case_when(
          #    Body == 0 ~ "very weak",
          #    Body == 1 ~ "weak",
          #    Body == 2 ~ "medium",
          #    Body == 3 ~ "strong",
          #    Body == 4 ~ "very strong"
          # ),
          # Body_chr = as.factor(Body_chr)
   # )


# model_body <- polr(factor(Body) ~ Smoky + Sweetness, data = whisky_df, Hess = TRUE)
# model_body <- polr(factor(Body) ~ Latitude, data = whisky_df, Hess = TRUE)
# model_body <- lm(Body ~ Longitude * Latitude, data = whisky_df)
# 
# summary(model_body)
# 
# model_body <- polr(Body ~ Smoky + Sweetness, data = model_data, Hess = TRUE)
# model_body <- polr(Smoky ~ Floral + Sweetness, data = model_data)
# 
# for (i in model_data$Longitude) {
#    print(i)
# }

# model_body <- lm(Body ~ Longitude + Latitude, data = whisky_df)





```


# Data visualization

## Longitude -- Latitude

```{r, eval=FALSE, include=FALSE}
theme_set(theme_linedraw())

# custom_palette <- RColorBrewer::brewer.pal(n = 10, "Blues")[5:10]

plot_long_function <- function(taste_var, taste_value, df) {
   
   df <- df %>% 
      filter(taste_var == taste_value)

   ggplot(df, aes(x = Longitude, y = Latitude, color = as.factor(!!enquo(taste_var)))) +
      geom_point(size = 5, alpha = 0.75) +
      # scale_color_ordinal(direction = -1, name = enquo(taste_var)) +
      scale_color_ordinal(direction = -1) +
      ggtitle(enquo(taste_value)) +
      geom_text_repel(aes(label = Distillery), show.legend = FALSE) 
   
}

plot_long_function(taste_category, "Smoky", whisky_df_long)


plot_wide_function <- function(taste_var, df) {
   
   df_select <- df %>% 
      select(!!enquo(taste_var), Longitude, Latitude, Distillery)

   ggplot(df_select, aes(x = Longitude, y = Latitude, color = as.factor(!!enquo(taste_var)))) +
      geom_point(size = 5, alpha = 0.75) +
      scale_color_ordinal(direction = -1, name = enquo(taste_var)) +
      ggtitle(enquo(taste_var)) +
      geom_text_repel(aes(label = Distillery), show.legend = FALSE) 
   
}

map(whisky_df %>% select(Body:Sweetness) , plot_wide_function, whisky_df)

plot_wide_function(Smoky, whisky_df)

plot_ll_function("Smoky", whisky_df)

map(names_taste, plot_ll_function, whisky_df)

map(whisky_df %>% select(Body:Sweetness), plot_ll_function, whisky_df)
lapply(whisky_df %>% select(Body:Sweetness), plot_ll_function, whisky_df)
map(whisky_df[3:4], plot_ll_function, whisky_df)

whisky_df %>% 
   select(Body:Sweetness) %>% 
   names() %>% 
   map(~ ggplot(whisky_df, aes_string(x = Longitude, y = Latitude, color = as.factor(.))) +
      geom_point(size = 5, alpha = 0.75) +
      scale_color_ordinal(direction = -1, name = .) +
      geom_text_repel(aes(label = Distillery), show.legend = FALSE))

whisky_df_long %>% 
   split(.$taste_category) %>% 
   map(~ggplot(., aes(x = Longitude, y = Latitude, color = as.factor(levels(taste_category)))) + 
          geom_point(size = 5, alpha = 0.75) +
          scale_color_ordinal(direction = -1, name = .) 
   )
   

```



