---
title: "Whisky Data"
subtitle: "Information about 86 Scottish Single Malt Whiskies and Distilleries"
output:
  html_document:
    theme: cosmo
    code_folding: show
    toc: true
    toc_float: true
---

<style type="text/css">
body, td {
   font-family: EB Garamond;
   font-size: 18px
}
h1.title {
   font-family: EB Garamond Medium;
   font-size: 38px
}
h1 { /* Header 1 */
   font-family: EB Garamond Medium;
   font-size: 28px
}
h2 { /* Header 2 */
   font-family: EB Garamond Medium;
   font-size: 22px
}
h3 { /* Header 3 */
   font-family: EB Garamond Medium;
   font-size: 18px
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(error=FALSE,
                      message=FALSE, 
                      warning=FALSE,
                      fig.width=7,
                      fig.height=5)
```

# R packages and setup

```{r packages}
# general (tidy-)toolkit
library(tidyverse)
library(stringr)

# displaying data
library(kableExtra)
library(skimr)
library(DT)

# statistical applications
library(broom)
library(MASS)
library(corrr)

# ggplot extensions
library(ggrepel)
library(ggdendro)
library(lemon)

# maps and co.
library(sf)
library(rnaturalearth)
library(tmap)
```

```{r ggplot setup, include=FALSE}
theme_set(theme_classic(base_family = "EB Garamond"))

theme_update(text = element_text(family = "EB Garamond"))

update_geom_defaults("text", list(family = theme_get()$text$family))
update_geom_defaults("label", list(family = theme_get()$text$family))
```

# Obtain data

Just download the file *whiskies.txt* `if` it is non-existing in the local path.

```{r}
whisky_data <- "whiskies.txt"

if (! whisky_data %in% list.files()) {
   
   whisky_url <- "https://outreach.mathstat.strath.ac.uk/outreach/nessie/datasets/whiskies.txt"
   
   download.file(whisky_url, whisky_data)
   
}

whisky_data_raw <- read.delim("whiskies.txt", sep = ",")
```

# Data management and wrangling {.tabset .tabset-fade}

Using different functions to get basic insights of the used whisky data.

```{r}
whisky_df <- whisky_data_raw %>%
   dplyr::select(! RowID) %>% 
   mutate(Postcode = str_trim(Postcode, side = "left")) 

# data in long format
whisky_df_long <- whisky_df %>% 
   pivot_longer(cols = Body:Floral, names_to = "taste_category", values_to = "value")

# summarising taste data
whisky_df_long %>% 
   group_by(taste_category) %>% 
   summarise(Mean = mean(value),
             Median = median(value),
             SD = sd(value)) %>% 
   rename(`Taste Category` = 1) %>% 
   kbl() %>% 
   kable_styling(html_font = "EB Garamond")
```

## `glimpse()`

```{r}
whisky_df %>% 
  glimpse()
```

## `head()`

```{r}
whisky_df %>% 
   head()
```

## `skim()`

```{r}
whisky_df %>%
   skim()
```
   
## `datatable()`

```{r}
whisky_df %>%
   datatable()
```

# Statistical analysis {.tabset .tabset-pills}

## Correlations

In order to analyze the relationship between the different taste categories the individual correlations (pearson $\rho$) are computed and displayed.

$$ \Large \rho = \frac{cov(X,Y)}{\sigma_X \sigma_Y}$$
```{r}
# Correlation matrix
whisky_df %>% 
   dplyr::select(Body:Floral) %>% 
   correlate() %>% 
   rearrange() %>%
   shave() %>% 
   fashion() 
   
# Correlation plot
corr_plot <- whisky_df %>% 
   dplyr::select(Body:Floral) %>% 
   correlate() %>% 
   rearrange() %>%
   shave() %>% 
   rplot() +
   theme(text = element_text(family = "EB Garamond"))


corr_plot 

ggsave("corr_plot.png", 
       corr_plot, 
       width = 8, 
       height = 4)

# Correlation network
network_plot <- whisky_df %>% 
   dplyr::select(Body:Floral) %>% 
   correlate() %>% 
   network_plot() + 
   theme(text = element_text(family = "EB Garamond"))

network_plot

ggsave("network_plot.png", 
       network_plot, 
       width = 8, 
       height = 4)

```

## Dendrogram {.tabset .tabset-fade}

### Dendrogram for distilleries

```{r, fig.height = 10, fig.width=6}
# Dendrogram distilleries
whisky_df %>% 
   column_to_rownames("Distillery") %>% 
   dplyr::select(Body:Floral) %>% 
   dist() %>% 
   hclust() %>% 
   ggdendrogram(rotate = TRUE) + 
   labs(title = "Dendrogram | Distilleries") + 
   theme(text = element_text(family = "EB Garamond"),
         plot.title = element_text(size = 20)) +
   ggsave("dendrogram_dist.png", 
          height = 10,
          width = 6)
```

### Dendrogram for taste categories

```{r, fig.height = 10, fig.width=6}
# Dendrogram taste categories
whisky_df %>% 
   dplyr::select(Distillery, Body:Floral) %>% 
   pivot_longer(cols = Body:Floral, names_to = "taste_category", values_to = "taste_value") %>% 
   pivot_wider(names_from = Distillery, values_from = taste_value) %>% 
   column_to_rownames("taste_category") %>% 
   dist() %>% 
   hclust() %>% 
   ggdendrogram(rotate = TRUE) +
   labs(title = "Dendrogram | Taste Categories") + 
   theme(text = element_text(family = "EB Garamond"),
         plot.title = element_text(size = 20)) +
   ggsave("dendrogram_taste.png", 
          height = 7.5,
          width = 4.5)
```

## *k*-means clustering

The *k*-means algorithm is used to create distinct clusters among the distilleries based on the values of the tasting categories. The goal of the *k*-means algorithm is to minimize the sum of within-cluster variance.

$$ \Large Formel$$

```{r}
set.seed(42)

# get k-means object for k from 2 to 9
for (i in 2:9) {
   
  assign(
     paste0("km_", i), 
     whisky_df %>% 
        dplyr::select(Body:Floral) %>% 
        kmeans(centers = i)
  )
   
}

# store in list
km_list <- mget(ls(pattern = "km_"))

# convert to tidy list with broom::tidy()
km_tidy_list <- km_list %>%
   map(tidy) 

# store in data frame / tibble
km_df <- km_tidy_list %>% 
   map(~ mutate(., cluster_max = max(as.numeric(cluster)))) %>% 
   reduce(full_join)

# function: check if row number is correct
gauss_sum_check <- function(n_max, n_start) {
   
   sum_1 <- (n_max * (n_max + 1)) / 2
   
   sum_2 <- (n_start * (n_start + 1)) / 2
   
   return(sum_1 - sum_2)
   
}

km_cluster_max <- km_df %>% 
   .$cluster_max %>% 
   max() # 9

# final check if k-means data frame has correct row number
gauss_sum_check(km_cluster_max, 1) == nrow(km_df)

# visualize k-means results
km_withinss <- km_df %>% 
   group_by(cluster_max) %>% 
   summarise(total_withinss = sum(withinss))

ggplot(km_withinss, aes(x = cluster_max, y = total_withinss)) + 
   geom_point() + 
   geom_line() + 
   scale_x_continuous(breaks = seq(2, 9, 1))
```

# Visualizing geographical information

```{r, fig.width = 7, fig.height = 10}
#  different background color for maps
theme_update(panel.background = element_rect(fill = "aliceblue"),
             plot.margin = margin(0, 0, 1.5, 0, "cm"))

british_isles <- ne_countries(country = c("united kingdom", "ireland"), 
                              returnclass = "sf",
                              scale = "medium")

# convert geographical whisky information
whisky_sf <- whisky_df %>% 
   st_as_sf(coords = c("Latitude", "Longitude")) %>%
   st_set_crs(27700) %>%
   st_transform(crs = 4326) %>% 
   mutate(lat = st_coordinates(.)[, 1],
          long = st_coordinates(.)[, 2])

# distilleries known for smoky whisky -- ENG
ggplot() + 
   geom_sf(data = british_isles, fill = "antiquewhite3") +
   geom_sf(data = subset(whisky_sf, Smoky == 4), alpha = 0.5) + 
   geom_text_repel(data = subset(whisky_sf, Smoky == 4), 
                   aes(label = Distillery, x = lat, y = long), 
                   family = "EB Garamond",
                   nudge_x = -1,
                   nudge_y = 0.125,
                   size = 5) +
   annotate("label", x = - 5.75, y = 60, label = "Distilleries with \n Smoky Whiskies", size = 8, fill = "antiquewhite") +
   coord_sf(x = c(-8, 0),
            y = c(53, 61)) +
   labs(x = "", y = "") + 
   ggsave("smoky_distilleries.png",
          width = 7,
          height = 10.5)

# distilleries known for non-smoky whisky -- ENG
ggplot() + 
   geom_sf(data = british_isles, fill = "antiquewhite3") +
   geom_sf(data = subset(whisky_sf, Smoky == 0), alpha = 0.5) + 
   geom_text_repel(data = subset(whisky_sf, Smoky == 0), 
                   aes(label = Distillery, x = lat, y = long), 
                   family = "EB Garamond",
                   nudge_x = 0.125,
                   nudge_y = -0.125,
                   size = 5) +
   annotate("label", x = - 5.75, y = 60, label = "Distilleries with \n Non-Smoky Whiskies", size = 8, fill = "antiquewhite") +
   coord_sf(x = c(-8, 0),
            y = c(53, 61)) +
   labs(x = "", y = "") + 
   ggsave("non-smoky_distilleries.png",
          width = 7,
          height = 10.5)

# distilleries known for sweet whisky -- ENG
ggplot() + 
   geom_sf(data = british_isles, fill = "antiquewhite3") +
   geom_sf(data = subset(whisky_sf, Sweetness == 4), alpha = 0.5) + 
   geom_text_repel(data = subset(whisky_sf, Sweetness == 4), 
                   aes(label = Distillery, x = lat, y = long), 
                   family = "EB Garamond",
                   nudge_x = -1,
                   nudge_y = 0.125,
                   size = 5) +
   annotate("label", x = - 5.75, y = 60, label = "Distilleries with \n Sweet Whiskies", size = 8, fill = "antiquewhite") +
   coord_sf(x = c(-8, 0),
            y = c(53, 61)) +
   labs(x = "", y = "") + 
   ggsave("sweet_distilleries.png",
          width = 7,
          height = 10.5)

# distilleries known for non-sweet whisky -- ENG
ggplot() + 
   geom_sf(data = british_isles, fill = "antiquewhite3") +
   geom_sf(data = subset(whisky_sf, Sweetness == 1), alpha = 0.5) + 
   geom_text_repel(data = subset(whisky_sf, Sweetness == 1), 
                   aes(label = Distillery, x = lat, y = long), 
                   family = "EB Garamond",
                   nudge_x = 0.125,
                   nudge_y = -0.125,
                   size = 5) +
   annotate("label", x = - 5.75, y = 60, label = "Distilleries with \n Non-Sweet Whiskies", size = 8, fill = "antiquewhite") +
   coord_sf(x = c(-8, 0),
            y = c(53, 61)) +
   labs(x = "", y = "") + 
   ggsave("non-sweet_distilleries.png",
          width = 7,
          height = 10.5)
```


































```{r old code, eval=FALSE, include=FALSE}
# model_data <- whisky_df %>% 
#    transmute(across(c(Body:Floral), as.factor))






whisky_df %>% 
   dplyr::select(Body:Floral)%>% 
   map(~ lm(Body ~ .x, data = whisky_df)) %>% 
   map(summary) %>% 
   map_dbl("r.squared") %>% 
   tidy %>% 
   arrange(desc(x)) %>% 
   rename(r.squared = x) %>% 
   knitr::kable()


# model_data <- whisky_df %>% 
#    mutate(across(c(Body:Floral), as.factor)) %>% 
#    mutate(Longitude = as.numeric(str_trim(as.character(Longitude)),
#           Latitude = as.numeric(str_trim(as.character(Latitude))))) %>% 
#    filter(is.infinite(Latitude))
   # mutate(Body = factor(Body, levels = 0:4, ordered = TRUE))
          # Body_chr = case_when(
          #    Body == 0 ~ "very weak",
          #    Body == 1 ~ "weak",
          #    Body == 2 ~ "medium",
          #    Body == 3 ~ "strong",
          #    Body == 4 ~ "very strong"
          # ),
          # Body_chr = as.factor(Body_chr)
   # )


# model_body <- polr(factor(Body) ~ Smoky + Sweetness, data = whisky_df, Hess = TRUE)
# model_body <- polr(factor(Body) ~ Latitude, data = whisky_df, Hess = TRUE)
# model_body <- lm(Body ~ Longitude * Latitude, data = whisky_df)
# 
# summary(model_body)
# 
# model_body <- polr(Body ~ Smoky + Sweetness, data = model_data, Hess = TRUE)
# model_body <- polr(Smoky ~ Floral + Sweetness, data = model_data)
# 
# for (i in model_data$Longitude) {
#    print(i)
# }

# model_body <- lm(Body ~ Longitude + Latitude, data = whisky_df)






theme_set(theme_linedraw())

# custom_palette <- RColorBrewer::brewer.pal(n = 10, "Blues")[5:10]

plot_long_function <- function(taste_var, taste_value, df) {
   
   df <- df %>% 
      filter(taste_var == taste_value)

   ggplot(df, aes(x = Longitude, y = Latitude, color = as.factor(!!enquo(taste_var)))) +
      geom_point(size = 5, alpha = 0.75) +
      # scale_color_ordinal(direction = -1, name = enquo(taste_var)) +
      scale_color_ordinal(direction = -1) +
      ggtitle(enquo(taste_value)) +
      geom_text_repel(aes(label = Distillery), show.legend = FALSE) 
   
}

plot_long_function(taste_category, "Smoky", whisky_df_long)


plot_wide_function <- function(taste_var, df) {
   
   df_select <- df %>% 
      select(!!enquo(taste_var), Longitude, Latitude, Distillery)

   ggplot(df_select, aes(x = Longitude, y = Latitude, color = as.factor(!!enquo(taste_var)))) +
      geom_point(size = 5, alpha = 0.75) +
      scale_color_ordinal(direction = -1, name = enquo(taste_var)) +
      ggtitle(enquo(taste_var)) +
      geom_text_repel(aes(label = Distillery), show.legend = FALSE) 
   
}

map(whisky_df %>% select(Body:Sweetness) , plot_wide_function, whisky_df)

plot_wide_function(Smoky, whisky_df)

plot_ll_function("Smoky", whisky_df)

map(names_taste, plot_ll_function, whisky_df)

map(whisky_df %>% select(Body:Sweetness), plot_ll_function, whisky_df)
lapply(whisky_df %>% select(Body:Sweetness), plot_ll_function, whisky_df)
map(whisky_df[3:4], plot_ll_function, whisky_df)

whisky_df %>% 
   select(Body:Sweetness) %>% 
   names() %>% 
   map(~ ggplot(whisky_df, aes_string(x = Longitude, y = Latitude, color = as.factor(.))) +
      geom_point(size = 5, alpha = 0.75) +
      scale_color_ordinal(direction = -1, name = .) +
      geom_text_repel(aes(label = Distillery), show.legend = FALSE))

whisky_df_long %>% 
   split(.$taste_category) %>% 
   map(~ggplot(., aes(x = Longitude, y = Latitude, color = as.factor(levels(taste_category)))) + 
          geom_point(size = 5, alpha = 0.75) +
          scale_color_ordinal(direction = -1, name = .) 
   )
   





# scotland_map <- map_data("world", "UK") %>% 
#    filter(subregion == "Scotland")
# 
# ggplot(scotland_map, aes(long, lat)) + 
#    geom_polygon() +
#    coord_quickmap()

   annotate("rect", xmin = - 8, xmax = -6, ymin = 59.5, ymax = 60.5, 
            alpha = 0.5, 
            fill = "antiquewhite",
            color = "antiquewhite4") +


```



